### 2022-04-14 DOC/methodology.tex

\subsection{Data description} \label{sec:data-description}

We first drew a random sample of 100,000 cases filed between 1st January 2014 and 31st December 2018 across India from the database published by Development Data Lab.\footcite{devdatalabs2021_eCourtsData} This window was chosen based on consultations with other researchers and practising advocates on the quality and completeness of data. We then attempted to download the case details for these 100,000 cases from the eCourts database.\footcite{ecourts2022} We were able to download case details for approximately 89,000 cases.\footnote{This difference was caused by non-functional links, and corrupted or non-machine data.} We used regular expressions to check whether the ``Act Name'' field references the \gls{ni}. This was the first level of case classification. However, from our previous experience with eCourts data, we expected more NI Act cases among cases where the Act Name field did not match our search parameters.\footcite[The Act Name field sometimes does not contain the name of the substantive act, but rather the name of the procedural Act (e.g., CrPC) or, in rare cases, the subject matter of the case. This is mainly because case classification and nomenclature practices vary from state-to-state. However, in rare cases, the field is not populated or contains erroneous information. For details, please see,][]{damle2020_ecourtsData} So we used regular expressions to search for references to the \gls{ni}, cheque bounce/dishonour along with \S~138 in the interim and final order texts, where available. We tagged the cases found through this protocol as related to cheque dishonour under \S~138 of the \gls{ni} (NI Act cases, for short). Data availability and quality vary significantly across States. Consequently, we could not include all the States of India in our study. We identified the States to exclude from the study based on the following criteria:

\begin{itemize}
\item States and Union Territories where we were unable to download sample data (this is because of data quality of the e-courts database);
\item States and Union Territories that do not classify any cases as relating to the Negotiable Instruments Act;
\item States and Union Territories where < 2\% of the final orders (as a proportion of total NI cases) are machine-readable and in English.
\end{itemize}

Based on the criteria above, of the 29 States (including Jammu and Kashmir) and 7 Union Territories, our final sample covers 8 States and 2 Union Territories, shown in Table \ref{tab:sample_desc}. For details concerning the selection, see appendix \ref{sec:sample_selection}.

{\footnotesize \begin{longtable}{@{}lrrr|r@{}}
\caption{Sample description}
\label{tab:sample_desc}\\
\toprule
\textbf{State} & \textbf{\gls{ni} cases} & \textbf{Non-\gls{ni} cases} & \textbf{\%\gls{ni} cases} & \textbf{Total cases}\\ \midrule
\endhead
Andhra Pradesh & 2638 & 18567 & 12.4 & 21205\\
Chandigarh & 731 & 1364 & 34.9 & 2095\\
Delhi & 5202 & 14742 & 26.1 & 19944\\
Goa & 399 & 2713 & 12.8 & 3112\\
Gujarat & 6752 & 49476 & 12.0 & 56228\\
Haryana & 5319 & 33542 & 13.7 & 38861\\
Himachal Pradesh & 1164 & 11948 & 8.9 & 13112\\
Karnataka & 11184 & 75807 & 12.9 & 86991\\
Maharashtra & 8875 & 82673 & 9.7 & 91548\\
Punjab & 5885 & 24739 & 19.2 & 30624\\
\midrule
\textbf{Total} & \textbf{48149} & \textbf{315571} & \textbf{13.2} & \textbf{363720}\\ \bottomrule
\end{longtable}
}

For these 10 States and Union Territories, we drew another random sample of 500,000 cases filed between 1st January 2014 and 31st December 2018 from the Development Data Labs database.\footcite{devdatalabs2021_eCourtsData} We then attempted to download case details for these 500,000 cases. We were able to successfully download case details for 4,68,855 cases. We removed cases from District and Sessions courts since these courts hear appeals and not the original matters. This gave a sample of 3,63,720 cases. As with the preliminary analysis, we used pattern-matching based on regular expressions to identify cases related to \gls{ni}. In total, our sample has 48,149 \gls{ni} cases. We manually checked the NI Act classification for approximately 1700 cases from the sample of 3,63,720 to check for erroneous classification. We found 1 false positive and no false negatives.\footnote{The false positive was because, for one case in Haryana, the court had dumped all the judgments issued on that date of disposal in the PDF file containing the order. As a result, the text for an actual NI Act case (along with all other cases disposed on that day) was erroneously included in the PDF of the judgment of that case. We further checked for similar erroneous data and found one more instance from the same court. However, none of the judgments in the PDF nor the case itself was related to NI Act. So effectively, it was correctly classified.} However, the data limitation above also limits our ability to reliably check for false negatives. Table \ref{tab:sample_desc} shows a summary of the sample. As shown, check dishonour cases in India represent approximately 13.2\% of courts' workload (pending and disposed). This number could be underestimated because orders cannot be parsed in approximately 36\% of the cases in our sample. Some may be NI Act cases, even though the Act Name field does not say so.

Table \ref{tab:state_disposal} shows the number of disposed and pending cases state-wise and the median time taken to dispose cases. According to \S~143(3) of the \gls{ni}, courts are required to, as far as reasonably possible, complete NI Act cases within six months, i.e. 180 days. However, as the table shows, the median time taken to dispose these cases is above the recommended limit in all states. Of the 37,479 disposed cases in the sample, 10,171 --- i.e. 27\% cases --- were disposed within the prescribed limit of six months.

{\footnotesize \begin{longtable}{@{}lrrrr@{}}
 \caption{State-wise disposed cases and time taken to dispose}\label{tab:state_disposal}\\
\toprule
\textbf{State} & \multicolumn{1}{p{2cm}}{\textbf{Disposed cases}} & \multicolumn{1}{p{2cm}}{\textbf{Pending cases}} & \multicolumn{1}{p{3cm}}{\textbf{Median days to dispose}} & \multicolumn{1}{p{3cm}}{\textbf{Median hearings to dispose}} \\
\midrule
Andhra Pradesh & 2154 & 484 & 451 & 16 \\
Chandigarh & 703 & 28 & 453 & 9 \\
Delhi & 3910 & 1292 & 518 & 8 \\
Goa & 280 & 119 & 352 & 16 \\
Gujarat & 5183 & 1569 & 323 & 13 \\
Haryana & 4241 & 1078 & 488 & 12 \\
Himachal Pradesh & 673 & 491 & 547 & 16 \\
Karnataka & 10221 & 963 & 218 & 6 \\
Maharashtra & 5024 & 3851 & 520 & 19 \\
Punjab & 5080 & 805 & 440 & 13 \\
\midrule
Overall & 37469 & 10680 & 395 & 11 \\
\bottomrule
\end{longtable}}

Table \ref{tab:state_pending} shows the pending cases across States and the duration of the pendency. The largest chunk of pending cases (35\%) has been pending for 2 to 3 years. There is state-wise variation in the distribution of pending cases. So the recommendation of the Supreme Court that the respective High Courts should find solutions specifically suited for their States, has merits.

{\footnotesize
\begin{longtable}{@{}lrrrrrr@{}}
 \caption{State-wise pending cases by duration}\label{tab:state_pending}\\
\toprule
\multirow{2}{*}{\textbf{State}} & \multicolumn{6}{c}{\textbf{Number of pending cases by duration (in years)}} \\
\cmidrule{2-7}
 & (1, 2] & (2, 3] & (3, 4] & (4, 5] & (5, 6] & (6, 7] \\
\midrule
\endhead
Andhra Pradesh & 45 & 196 & 120 & 80 & 29 & 14 \\
Chandigarh & 0 & 6 & 16 & 5 & 1 & 0 \\
Delhi & 127 & 393 & 494 & 195 & 69 & 14 \\
Goa & 23 & 41 & 25 & 26 & 4 & 0 \\
Gujarat & 189 & 481 & 415 & 345 & 95 & 44 \\
Haryana & 201 & 578 & 204 & 82 & 9 & 4 \\
Himachal Pradesh & 47 & 142 & 120 & 96 & 56 & 28 \\
Karnataka & 134 & 422 & 253 & 97 & 40 & 17 \\
Maharashtra & 418 & 1122 & 921 & 787 & 431 & 172 \\
Punjab & 150 & 414 & 156 & 55 & 24 & 6 \\
\midrule
\textbf{Total} & 1334 & 3795 & 2724 & 1768 & 758 & 299 \\
\bottomrule
\end{longtable}
}

Table \ref{tab:yearWise_ni} shows the year-wise number of \gls{ni} cases in our sample. As expected, the pendency rate for older cases is lower than newer ones.

\begin{longtable}{@{}lrr|r@{}}
 \caption{Yearwise number of NI Act cases}\label{tab:yearWise_ni}\\
\toprule
\textbf{Year of filing} & \textbf{Disposed} & \textbf{Pending} & \textbf{Total} \\
\midrule\endhead
2014 & 6293 & 391 & 6684 \\
2015 & 6436 & 888 & 7324 \\
2016 & 8562 & 2072 & 10634 \\
2017 & 8904 & 3149 & 12053 \\
2018 & 7274 & 4180 & 11454 \\
\midrule
Total & 37469 & 10680 & 48149 \\
\bottomrule
\end{longtable}

%\subsection{Approach to analysis} \label{sec:approach-analysis}

Our objective is two-fold, to examine whether the interventions proposed by the Amicus Curiae are likely to affect the duration of cheque-dishonour cases, and if so, what is the magnitude of the effect. To that end, we first identified cases with specific characteristics that are thought to increase the duration of a cheque-dishonour case.

\subsection{Identifying case characteristics} \label{sec:text-mining}

To identify cases with the characteristics of interest, we analysed the interim and final orders using a pattern matching protocol and regular expressions implemented in Python 3.8. We supplemented this with the information reported in the case details downloaded from the eCourts database. In particular, we relied on the purpose of hearings and the nature of disposal. The characteristics and the strategy to identify whether they appear in a given case are as follows:

\begin{description}
\item [Non-appearance by the accused] -- We looked for references to \S~72 \gls{crpc},\footnote{\S~72, Code of Criminal Procedure. Warrants to whom directed.} and terms like - accused not available / out of station / absconding / not present / absent, or that accused’s presence could not be secured, or references to the issuance of a bailable / non-bailable warrant to secure the presence of the accused. In addition to this, we looked at references to non-appearance, absconsion, bailable / non-bailable warrants in the data on the purpose of hearings.

Table \ref{tab:nonAppearance_yearWise} shows the number of cases in the sample in which the accused failed to appear in at least one of the hearings. It appears that this is a widespread occurrence, and consistently so, over the entire observation period.

\begin{longtable}{@{}lrrr@{}}
 \caption{Cases where the accused was absent for at least one hearing}\label{tab:nonAppearance_yearWise}\\
\toprule
\multirow{2}{*}{Year} & \multirow{2}{*}{Total NI Act cases} & \multicolumn{2}{c}{Non Appearance}\\
\cmidrule{3-4}
&& Number of cases & Percentage \\
\midrule\endhead
2014 & 6684 & 4302 & 64.4 \\
2015 & 7324 & 5241 & 71.6 \\
2016 & 10634 & 7527 & 70.8 \\
2017 & 12053 & 8628 & 71.6 \\
2018 & 11454 & 7927 & 69.2 \\
\bottomrule
\end{longtable}

We expected cases with incidences of absence by the accused would take more days and more hearings to dispose.

\item[]

\item [Jurisdiction issue] -- We looked for references to lack of jurisdiction / lack of proper jurisdiction / beyond jurisdiction, references to \S~142 of the \gls{ni},\footnote{\S~142, Negotiable Instruments Act. Cognisance of offences.} and citations of specific cases like \citetitle{sc2014_dhanuka},\footcite{sc2014_dhanuka} and \citetitle{sc2016_carbon}.\footcite{sc2016_carbon} Additionally, we looked at cases that were marked transferred from one court to another or were observed to be transferred in the hearings or disposal data, since a vast majority of such transfers are due to jurisdictional issues.

Table \ref{tab:jurisdiction_yearWise} shows the number of cases in the sample in which there was a jurisdictional issue. The proportion of such cases as a percentage of the total, though significant, has fallen slightly over the observation period. The increase in cases with jurisdictional issues in 2016 is likely to be a consequence of the amendment in 2015, which resulted in more cases being transferred to courts that would now have appropriate jurisdiction.

\begin{longtable}{@{}lrrr@{}}
 \caption{Cases with jurisdictional issues}\label{tab:jurisdiction_yearWise}\\
 \toprule
 \multirow{2}{*}{Year} & \multirow{2}{*}{Total NI Act cases} & \multicolumn{2}{c}{Jurisdictional issue}\\
 \cmidrule{3-4}
 && Number of cases & Percentage \\
 \midrule\endhead
 2014 & 6684 & 2089 & 31.3 \\
 2015 & 7324 & 2229 & 30.4 \\
 2016 & 10634 & 3520 & 33.1 \\
 2017 & 12053 & 3360 & 27.9 \\
 2018 & 11454 & 2900 & 25.3 \\
 \bottomrule
\end{longtable}

We expected cases with jurisdictional issues to take more days and more hearings to dispose.

\item[]

\item [Case being converted to a summons trial] -- We looked for references to the case being tried under \S~143 of the NI Act,\footnote{\S~143, Negotiable Instruments Act. Power of Court to try cases summarily.} or \S\S~262, 263, 264 or 265 of the \gls{crpc},\footnote{\S\S~262, 263, 264 or 265, Code of Criminal Procedure. Procedure for summary trials; Record in summary trials; Judgement in cases tried summarily; Language of record and judgment.} or references to the case being tried summarily. These cases were marked as not summons trials. Cases where these were deemed not to apply, or the case was being conducted normally, or it is undesirable to try the case summarily, or the sections above not being applicable were marked as summons trials. Further, cases, where any hearing mentioned summons trial as the purpose, were marked as summons cases.

Table \ref{tab:summons_yearWise} shows the number of cases in the sample in which the case was converted from a summary trial to a summons trial. The proportion of such cases as a percentage of the total has grown over the observation period.

\begin{longtable}{@{}lrrr@{}}
 \caption{Cases converted to summons trials}\label{tab:summons_yearWise}\\
 \toprule
 \multirow{2}{*}{Year} & \multirow{2}{*}{Total NI Act cases} & \multicolumn{2}{c}{Summons trial}\\
 \cmidrule{3-4}
 && Number of cases & Percentage \\
 \midrule\endhead
 2014 & 6684 & 1251 & 18.7 \\
 2015 & 7324 & 1880 & 25.7 \\
 2016 & 10634 & 2814 & 26.5 \\
 2017 & 12053 & 3542 & 29.4 \\
 2018 & 11454 & 3389 & 29.6 \\
 \bottomrule
\end{longtable}

We expected cases converted to summons trial to take more days and more hearings to dispose.

\item [Case referred to mediation] -- We looked for references to the case being referred to mediation / Lok Adalat / \gls{adr}, or references to the matter being settled mutually, or compromised. We also looked at references to the cases being sent to mediation in the hearings data, and references to the case concluding in the Lok Adalat or being settled mutually, or the parties reaching a compromise.

Table \ref{tab:mediation_yearWise} shows the number of cases in the sample in which the accused failed to appear in at least one of the hearings. The proportion of such cases is considerable over the entire observation period.

\begin{longtable}{@{}lrrr@{}}
 \caption{Cases referred to mediation}\label{tab:mediation_yearWise}\\
 \toprule
 \multirow{2}{*}{Year} & \multirow{2}{*}{Total NI Act cases} & \multicolumn{2}{c}{Mediation}\\
 \cmidrule{3-4}
 && Number of cases & Percentage \\
 \midrule\endhead
 2014 & 6684 & 997 & 14.9 \\
 2015 & 7324 & 1521 & 20.8 \\
 2016 & 10634 & 2609 & 24.5 \\
 2017 & 12053 & 3100 & 25.7 \\
 2018 & 11454 & 2484 & 21.7 \\
 \bottomrule
\end{longtable}

\item[]

\item [Multiplicity of proceedings] -- We looked for references to multiple cheques being dishonoured and multiple transactions, and references to \S\S~219 and 220 of the \gls{crpc}.\footnote{\S\S~219 and 220, Code of Criminal Procedure. Three offences of same kind within year may be charged together; Trial for more than one offence.}

Table \ref{tab:mult_yearWise} shows the number of cases in the sample in which there was a multiplicity of proceedings. As a percentage of the total, the proportion of such cases is small. However, as \Cref{tab:state_multiplicity} shows, the burden caused due to multiplicity varies across States.

\begin{longtable}{@{}lrrr@{}}
 \caption{Cases with a multiplicity of proceedings}\label{tab:mult_yearWise}\\
\toprule
\multirow{2}{*}{Year} & \multirow{2}{*}{Total NI Act cases} & \multicolumn{2}{c}{Multiplicity of proceedings}\\
\cmidrule{3-4}
&& Number of cases & Percentage \\
\midrule\endhead
2014 & 6684 & 259 & 3.9 \\
2015 & 7324 & 382 & 5.2 \\
2016 & 10634 & 544 & 5.1 \\
2017 & 12053 & 515 & 4.3 \\
2018 & 11454 & 310 & 2.7 \\
\bottomrule
\end{longtable}

We expected cases with a multiplicity of proceedings to take more days and more hearings to dispose.

\item[]

\item [Contested cases] -- The fact of the case being contested or uncontested is reported by eCourts in the ``disposal type'' field. We take the information as reported.

  Table \ref{tab:mult_yearWise} shows the number of cases in the sample which were contested by the accused. The proportion of such cases is significant but appears to have declined over the years. This may be due to an increased tendency of parties to compound a case.\footnote{\S~147 of the Negotiable Instruments Act provides that every offence punishable under the Act shall be compoundable. \S~147, Negotiable Instruments Act. Offences to be compoundable.}

\begin{longtable}{@{}lrrr@{}}
 \caption{Contested cases}\label{tab:cont_yearWise}\\
\toprule
\multirow{2}{*}{Year} & \multirow{2}{*}{Total NI Act cases} & \multicolumn{2}{c}{Contested cases}\\
\cmidrule{3-4}
&& Number of cases & Percentage \\
\midrule\endhead
2014 & 6684 & 2171 & 32.5 \\
2015 & 7324 & 1944 & 26.5 \\
2016 & 10634 & 1822 & 17.1 \\
2017 & 12053 & 1478 & 12.3 \\
2018 & 11454 & 868 & 7.6 \\
\bottomrule
\end{longtable}

We expected contested cases to take more days and more hearings to dispose than uncontested cases.

\item[]

\end{description}

These case characteristics were represented as binary variables. For example, if the parties did not appear for a hearing in a given case, the Non-appearance variable takes the value 1. If the parties appeared at every hearing, we would mark it as 0. Similarly, for the Summons Trial variable, if the case was converted to a summons trial, it is marked as 1. Otherwise, it is marked as 0.

%\subsection{Overview of the data} \label{sec:overview-data}

Table \ref{tab:case_chars} shows the State-wise number of cases in which we were able to identify characteristics of interest. The row totals are not expected to add up to the total cases in the respective state in the sample, because some cases can have more than one of the selected characteristics. As an illustration, a case may involve the dishonour of multiple cheques and be referred to mediation.

{\footnotesize
 \begin{longtable}{@{}lrrrrrr@{}}
 \caption{Cases with identified characteristics of interest}
 \label{tab:case_chars}\\
 \toprule
 \textbf{State} & \textbf{Non} & \textbf{Summons} & \textbf{Mediation} & \textbf{Jurisdiction} & \textbf{Multiplicity} & \textbf{Contested} \\
 & \textbf{Appearance} & \textbf{Trial} & & \textbf{Issue} & &\\
 \midrule
 \endhead
 Andhra Pradesh & 1674 & 814 & 260 & 210 & 124 & 753 \\
 Chandigarh & 701 & 408 & 204 & 278 & 53 & 106 \\
 Delhi & 2062 & 1359 & 1026 & 1045 & 208 & 521 \\
 Goa & 381 & 78 & 90 & 33 & 18 & 109 \\
 Gujarat & 5125 & 391 & 402 & 3059 & 107 & 797 \\
 Haryana & 5214 & 3376 & 1731 & 2109 & 540 & 599 \\
 Himachal Pradesh & 864 & 351 & 506 & 299 & 33 & 113 \\
 Karnataka & 5850 & 2002 & 1846 & 953 & 410 & 3615 \\
 Maharashtra & 5933 & 963 & 2546 & 3831 & 135 & 1118 \\
 Punjab & 5821 & 3134 & 2100 & 2281 & 382 & 552 \\
 \bottomrule
\end{longtable}
}

The State-wise numbers presented in \Cref{tab:case_chars} are likely underestimated. This is because orders cannot be parsed in approximately 36\% of the cases in our sample. Further, some courts do not adequately record details of the purpose of hearings and the nature of disposal \autocite{damle2020_ecourtsData}. Cases in which information regarding these characteristics could not be found were marked as not having those particular characteristics. In our manual check of approximately 700 cases, we did not find any false positives or false negatives. However, the data limitation above also limits our ability to reliably check for false negatives. This, in turn, means there could be many more cases that have the characteristics above but could not be identified as having them.

\subsubsection{Regression model specifications} \label{sec:model-selection}
\input{model}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper_chequeDishonour"
%%% End: